---
title: "Programmierung R - Exercise"
author: "Data manipulation"
date: "April 12, SS 2022 || Hannah Behrens"

toc: true
tocoverview: false
tocheader: Inhaltsübersicht
titlefontsize: 22pt

bibliography: referencescswr.bib
biblio-style: authoryear-comp

output: fhswf::presentation
knit: rmarkdown::render

nocite: |
  @cswr
---

```{r setup, include=FALSE}
library(knitr)
options(digits = 4)
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, cache=TRUE, 
                      dev.args=list(pointsize=11), size="scriptsize", fig.height = 4, attr.source='.numberLines')

# Packages
library(car)
library(fGarch)
library(gtools)
library(HistogramTools)
library(ineq)
library(kableExtra)
library(texreg)
library(tidyverse)
library(vcd)
library(latex2exp)
library(knitcitations)
library(gridExtra)
library(plotly)
library(rgl)
library(mgcv)
```

## Task 1

### Your turn

In your R environment the data set `datasets::airquality` [@datasets] is already available.

1.1 Make yourself familiar with this data set by examining

- its data type

- its structure (and dimensions) and

- its variables.

1.2 Where do and how many NAs occur in the data set? Remove those observations (full rows) and save this data set as `airquality2`. How many rows have been deleted?


## Task 1.1 - answer

1.1 Make yourself familiar with the data set `datasets::airquality` [@datasets]...

```{r}
typeof(airquality) # its data type

str(airquality) # its structure and its variables

?airquality # description of the data set
```
$\rightarrow$ This data set is a list, more precisely a data frame.

$\rightarrow$ 153 observations (153 rows) and 6 variables (6 columns)

## Task 1.2 - answer

$\rightarrow$ 37 NAs are present in the column `Ozone` and 7 NAs in `Solar.R`.

```{r}
summary(airquality) # to see where and how many NAs occur
```

\begin{textblock}{6}(6,6)
\begin{alertblock}{}
Which return value would you expect if \texttt{summary()} were applied to a categorical variable?
\end{alertblock}
\end{textblock}

## Task 1.2 - answer

```{r}
# remove those observations (full rows), where NAs occur:
airquality2 <- na.omit(airquality) 
nrow(airquality) - nrow(airquality2) # 42 rows have been removed 
# consequently, 153 - 42 = 111 rows left
```

## Task 2

### Your turn

As we have already seen, the variable `Month` is numeric.

2.1) Add a further variable named `Month_name` that saves the name of the month (as character) to the *NAs-free* `airquality2` data set.

2.2) Apply `summary()` to `Month_name` and check your previous assumption. Assign the result to the variable `summary_air`.


## Task 2.1 - answer

As we have already seen, the variable `Month` is numeric.

Add a further variable named `Month_name` that saves the name of the month (as character) to the *NAs-free* `airquality2` data set.

```{r}
levels_month <- unique(airquality2$Month) # get unique values of Month
# add the names of the months as a factor to the data set:
airquality2$Month_name <- factor(x = airquality2$Month,
                               levels = levels_month, # alternatively: c(5, 6, 7, 8, 9)
                               # labels: how to name the levels, 
                               # make use of the predefined  constant named month.name
                                 labels = month.name[levels_month])

head(airquality2) # check the modification
```

## Task 2.2 - answer

```{r}
summary_air <- summary(airquality2$Month_name)
summary_air # a vector results

typeof(summary_air) 
```


## Task 3

### Your turn

\begin{alertblock}{}
Apply the following tasks to the modified \texttt{airquality2} data set. 
\end{alertblock}

The temperature of the `airquality` data set is given in degrees F (Fahrenheit). 

3.1 Convert these values into degrees Celsius [C] by using the following formula given by @NIST2021: $°C = \frac{(°F - 32)}{1.8}$.\newline

3.2 You are unhappy with the variables' names of the `airquality2` data set. Get the column names and change `Temp` to `Temperature`. 


## Tasks 3.1 and 3.2 - answers

3.1 

```{r}
# convert temperature from degrees Fahrenheit to degrees Celsius
# first option:
airquality2 <- transform(airquality2,
          Temp = (Temp - 32)/1.8)
# second option:
# airquality2$Temp <- (airquality2$Temp - 32)/1.8
```


3.2 

```{r}
colnames_airqu <- colnames(airquality2) # get column names of airquality2
colnames_airqu

colnames_airqu[4] <- "Temperature" # change "Temp" to "Temperature"
colnames_airqu # check the modified vector
```

## Task 3.2 - answer

```{r}
colnames(airquality2) <- colnames_airqu # assign modified column names to data set
colnames(airquality2) # check the modification
```

\begin{textblock}{4}(5,2)
\begin{alertblock}{}
Check your modifications!
\end{alertblock}
\end{textblock}


## Task 4

### Your turn

4.1 Imagine the following observation `new_obs` has misleadingly not been integrated in the `airquality` data set. Add this observation to the modified data set, assign it to the variable `airquality3` and check the dimensions of the data set afterwards.

```{r}
new_obs <- data.frame(Ozone = 20, Solar.R = 300, Wind = 8.7,
                      Temperature = 20, Month = 6, Day = 6, Month_name = "June")
```

4.2  Add a date column to the modified `airquality3` data set:

  - First paste the entries of the `Month` and `Day` column with the year of the observations by separating the single values by "/".
  
  - Then convert these strings to POSIX objects.
  
4.3 Save the modified `airquality3` data set as CSV-file. Afterwards, control your created CSV-file by reading it in your console.

## Tasks 4.1 and 4.2 - answers

4.1

```{r}
airquality3 <- rbind(airquality2, new_obs) # add observation new_obs to the
# modified airquality data set (airquality2) 
dim(airquality3) # check the dimensions (+ 1 row)
```
4.2 

```{r}
# paste Month, Day and year together in form of Month/Day/Year e.g. 5/2/1973
airquality3$Date <- paste(airquality3$Month, airquality3$Day, "1973", sep = "/")
# convert these strings to POSIX objects based on strptime()
airquality3$Date <- strptime(airquality3$Date, format = "%m/%d/%Y")

head(airquality3) # check the modifications
```

## Task 4.3 - answer

```{r, eval=FALSE}
# save the modified data set as a csv-file:
write.table(airquality3, file = "Airquality3.csv", sep = ",")
getwd() # get working directory
```
```{r}
airquality3_csv <- read.table(file = "Airquality3.csv", sep = ",") # read csv-file
head(airquality3_csv)
```

\begin{textblock}{8}(3,1)
\begin{alertblock}{}
We will use this modified data set in a future exercise.
\end{alertblock}
\end{textblock}

\begin{textblock}{8}(3,8)
\begin{alertblock}{}
DataCamp course to importing data in R will follow!
\end{alertblock}
\end{textblock}


## Task 5

### Your turn

Answer the following questions by using `airquality3`.

5.1 How many observations with temperatures above 30°C exist in the data set?

5.2 Sort the modified `airquality3` data set by temperature in descending order and only return the ordered values of the variable `Wind`.

5.3 Split `airquality3` into sub populations by `Month_name` and return the minimum of each variable. Does it work properly or are some changes necessary?

5.4 Return only observations of those months that include a "u" in their names.

## Tasks 5.1 and 5.2 - answers

5.1 

```{r}
sum(airquality3$Temperature > 30) # Temperatures above 30°C have been measured 20 times
```
5.2

```{r}
# sort by temperature in descending order and return the values of Wind:
airquality3$Wind[sort(airquality3$Temperature, decreasing = TRUE)] 
```

## Task 5.3 - answers

5.3
```{r, echo=TRUE, eval=FALSE}
aggregate(x = airquality3, by = list(Month_name = airquality3$Month_name), FUN = min) 
# error results
```

$\rightarrow$ It does not work since `Month_name` is an unordered factor, so order the levels of `Month_name` and apply `aggregate()` again.

```{r}
airquality3$Month_name <- factor(x = airquality3$Month_name, ordered = TRUE)
# remove seventh column, otherwise Month_name would be listed twice
aggregate(x = airquality3[,-7], by = list(Month_name = airquality3$Month_name), 
          FUN = min)
```


## Task 5.4 - answers

5.4

```{r}
# search for "u" in Month_name and only return the corresponding observations 
airquality4 <- airquality3[grep(pattern = "u", x = airquality3$Month_name),] 

head(airquality4)

dim(airquality4)
```

## Task 6

### Your turn

6.1 Create a data frame `month_seasons` with two columns: In the first column called `Month_name` the names of all (12) months are listed. The second column named `Season` consists of the seasons (Spring, Summer, Autumn, Winter) corresponding to the months.

6.2 Then, merge `month_seasons` and `airquality3` by `Month_name` and save this data frame as `airquality3`. Check the dimensions of the resulting data set.


## Task 6.1 - answer

6.1 
```{r}
month_seasons <- data.frame(Month_name = factor(x = month.name, levels = month.name, 
                            ordered = TRUE), # order months
                            Seasons = factor(x = c("Winter", "Winter", "Spring", 
                            "Spring", "Spring", "Summer", "Summer", "Summer", "Autumn", 
                            "Autumn", "Autumn", "Winter"), 
                            levels = c("Spring", "Summer", "Autumn", "Winter"),  
                            ordered = TRUE)) # order seasons

head(month_seasons, n = 7)
```

## Task 6.2 - answer

6.2

```{r}
airquality3 <- merge(x = airquality3, y = month_seasons, by = "Month_name")
dim(airquality3) # + 1 column since column Seasons has been added
```


## Task 7

**Do not** run the following code on your console!

```{r, eval=FALSE}
x <- c(1:10, 23:19, NA)
X <- 3
x <- x[x < 23]
x <- x[x <= 22]
x + X
x[X] <- x[X] + X
x <- x[X] + X
x <- x %% 3
y <- as.logical(x)
```

### Your turn

What is the result of

+ `x` after executing lines 1--6?

+ `x` after executing lines 1--8?

+ `y`?


## Task 7 - answer

What is the result of `x` after executing lines 1--6?

```{r}
#>R [1]  1  2  6  4  5  6  7  8  9 10 22 21 20 19 NA
```

What is the result of `x` after executing lines 1--8?

```{r}
#>R [1] 0
```

What is the result of `y`?

```{r}
#>R [1] FALSE
```


## Task 8

**Do not** run the following code on your console!

```{r}
L <- list(money = c(250, 124, 360, 720, 340, 340),
          hours = c(19, 12, 30, 48, 26, 25),
          idx = c(1:6),
          name = c("Paul", "Emma", "Mia", "John", "Kim"))
z <- 2
L[[4]][[6]] <- "Maxi"
L$hours <- L$hours * 1.3
L[[1]] <- L$money + 25
L[1] <- 1250
L[[4]][[2]] <- L[[4]][[6]]
L[[z]] <- 150
```

### Your turn

What is the result of

+ `L` after executing lines 1--8?

+ `L` after executing lines 1--9?

+ `L` after executing lines 1--11?


## Task 8 - answers

What is the result of `L` after executing lines 1--8?

```{r}
#>R 
#>R $money
#>R [1] 275 149 385 745 365 365

#>R $hours
#>R [1] 24.7 15.6 39.0 62.4 33.8 32.5

#>R $idx
#>R [1] 1 2 3 4 5 6

#>R $name
#>R [1] "Paul" "Emma" "Mia"  "John" "Kim"  "Maxi"

# since "Maxi" has been added to the name list,
# since each element of the hours list has been multiplied by 1.3 and
# since to each element of the money list 25 has been added
```

## Task 8 - answers

What is the result of `L` after executing lines 1--9?

```{r}
#>R $money
#>R [1] 1250

#>R $hours
#>R [1] 24.7 15.6 39.0 62.4 33.8 32.5

#>R $idx
#>R [1] 1 2 3 4 5 6

#>R $name
#>R [1] "Paul" "Emma" "Mia"  "John" "Kim"  "Maxi"

# in addition to the previous steps, the value 1250 has been assigned as the first list
# (money) of L
```

## Task 8 - answers

What is the result of `L` after executing lines 1--11?

```{r}
#>R $money
#>R [1] 1250

#>R $hours
#>R [1] 150

#>R $idx
#>R [1] 1 2 3 4 5 6

#>R $name
#>R [1] "Paul" "Maxi" "Mia"  "John" "Kim"  "Maxi"

# in addition to the previous steps, the second element of the name list has been 
# overwritten by the sixth name and
# the value 150 has been assigned as the second list (hours) of L
```


## References 







