---
title: "Programmierung R - Exercise"
author: "Software Development"
date: "May 3, SS 2022 || Hannah Behrens"

toc: true
tocoverview: false
tocheader: Inhaltsübersicht
titlefontsize: 22pt

bibliography: C:/Users/Mitarbeiterin/sciebo/nongit/SoSe_22/cswr_R_Programmierung_2022/Uebungen_Slides/referencescswr.bib
biblio-style: authoryear-comp

output: fhswf::presentation
knit: rmarkdown::render

nocite: |
  @shiny
---

```{r setup, include=FALSE}
library(knitr)
options(digits = 4)
knitr::opts_chunk$set(echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE, cache=TRUE, 
                      dev.args=list(pointsize=11), size="scriptsize", fig.height = 4, attr.source='.numberLines')

# Packages
library(car)
library(fGarch)
library(gtools)
library(HistogramTools)
library(ineq)
library(kableExtra)
library(texreg)
library(tidyverse)
library(vcd)
library(latex2exp)
library(knitcitations)
library(gridExtra)
library(plotly)
library(rgl)
library(mgcv)
library(readxl)
library(dplyr)
```

## Today's problem - tourism data

We will work with the Australian `tourism` data provided by @Hyndman2021: The file can be downloaded [**here**](https://bit.ly/fpptourism) or [**here**](http://robjhyndman.com/data/tourism.xlsx) and read into R with `readxl::read_excel()` [@readxl] (see announcement from May 2, 2022 in moodle).

The original tourism data was published by the [Tourism Research Division (Tourism Research Australia) of the Australian Trade and Investment Commission of the Australian Government](https://www.tra.gov.au/data-and-research).


## Task - Understanding the tourism data set

### Your turn

Make yourself familiar with the Australian tourism data set by making a sketch of how the variables (especially `State`, `Region` and `Purpose`) relate to each other.

## Understanding the tourism data set - answer

\begin{figure}
\includegraphics[width=11.5cm]{understanding_tourism_data.pdf}
\caption{Sketch of the variables of the Australian tourism data set based on Hyndman and Athanasopoulos (2021).}
\end{figure}

\begin{textblock}{11}(1,1)
\begin{alertblock}{}
See \url{https://otexts.com/fpp3/hts.html} by Hyndman and Athanasopoulos (2021)
\end{alertblock}
\end{textblock}


## Today's problem - tourism data

**Aim:**

We are interested in considering the overnight trips of a specific region in a specific state in Australia. Optionally, we want to select a specific purpose for the overnight trips, e.g. we are interested in the time-dependent visits of business people in Adelaide in South Australia.

On the one hand, we want to get and save the filtered data and on the other hand, we want to visualize the resulting time series in a nice ggplot.
Furthermore, we are also interested in the total number of overnight trips for each region in a state like South Australia.



## Task - Making a plan

### Your turn

Make a plan to solve the problem, i.e. make a plan for the implementation. What do you have to do? Which functions are needed? Make a sketch with some notes.


## Making a plan - answer

\begin{figure}
\includegraphics[width=12cm]{Plan_aus_tourism_data.pdf}
\caption{A plan for the implementation.}
\end{figure}

## Making a plan

Concretely,

- create a folder, e.g. named `aus_tourism`

- read in the data in a main program named `main_program_aus_tourism` and call the following functions:

    - `selectedData()`,

    - `plotTs()` and

    - `summarizedRegion()`
    
  which we have to write.


## Task - Coding

### Your turn

Write the corresponding program and its needed functions. For each R function create an own R file. Start with:

```{r}
library(ggplot2)
library(readxl)
library(dplyr)

getwd()
setwd("aus_tourism") # data set and functions are saved in folder "aus_tourism"
tourism <- read_excel("tourism.xlsx")
```


## Coding - selectedData() und plotTs() - answer

```{r}
selectedData <- function(dataset, state, region, purpose = NULL){
  
  filtered_Data <- dataset %>% 
    filter(State == state, Region == region) 
  
  if (!is.null(purpose)) { # filter either for a specific purpose or not
    filtered_Data <- filtered_Data %>%
      filter(Purpose == purpose)
  }
  return(filtered_Data)
}
```

```{r}
plotTs <- function(filtered_data){
    ts_plot <- ggplot(data = filtered_data, mapping = aes(x = as.Date(Quarter), 
                                                   y = Trips, 
                                                   color = Purpose))+ 
      geom_line()+
      xlab("Time [Quarter]")+
      ylab("Overnight trips ('000)")
    ts_plot # return the created ggplot
}
```

\begin{textblock}{5}(7, 3.5)
\begin{alertblock}{}
Make use of the R package \texttt{dplyr} (Wickham, François, et al. (2021)).
\end{alertblock}
\end{textblock}


## Coding - summarizeRegion() - answer

```{r}
summarizeRegion <- function(dataset, state){
  sum_R <- dataset %>% 
    filter(State == state) %>% 
    group_by(Region) %>% 
    summarize(Trips = sum(Trips))
  sum_R <- as.data.frame(sum_R)
  return(sum_R)
}
```

\begin{textblock}{5}(7, 1)
\begin{alertblock}{}
Make use of the R package \texttt{dplyr} (Wickham, François, et al. (2021)).
\end{alertblock}
\end{textblock}

## Coding - main program - answer

```{r, echo=FALSE, eval=TRUE}
gwd <- getwd()
setwd("aus_tourism") # data set and functions are saved in folder "aus_tourism"
tourism <- read_excel("tourism.xlsx")
dataset <- tourism
source("selectedData.R")
source("plotTs.R")
source("summarizeRegion.R")
```


```{r, eval=FALSE}
library(ggplot2)
library(readxl)
library(dplyr)

getwd()
setwd("aus_tourism") # data set and functions are saved in folder "aus_tourism"
tourism <- read_excel("tourism.xlsx")

source("selectedData.R")
source("plotTs.R")
source("summarizeRegion.R")

state <- "South Australia"
region <- "Adelaide" 
purpose <- "Holiday"
dataset <- tourism 

sel_Data <- selectedData(dataset = dataset, state = state, region = region)
plot1 <- plotTs(filtered_data = sel_Data)
sR <- summarizeRegion(dataset = dataset, state = state)
# save the filtered data:
write.table(sel_Data, file = "Filtered_Aus_tourism_data.csv", sep = ",")

png("Tourism_data_South_Aus_Adelaide.png") # additionally, save the plot
plot1
dev.off()
```


## Task - Documenting

### Your turn

Comment and document your program and your functions as it has been shown in the lecture by @cswr.

When documenting your functions, make use of the R package `roxygen2` [@roxygen2].

## Documenting - answer

See the files:

- `main_program_aus_tourism.R`:

- `selectedData.R`,

- `plotTs.R` and

- `summarizedRegion.R`

## Task - Testing

### Your turn

Test your program and functions by filtering 

1) South Australia as state, Adelaide as region and **do not** select a specific purpose. Plot the filtered time series.

2) South Australia as state, Adelaide as region and select Holiday as purpose. Plot the filtered time series.

Do some more tests to ensure that your functions work appropriately.


## Testing 1) - answer - South Australia as state and Adelaide as region


```{r, eval=TRUE}
state <- "South Australia"
region <- "Adelaide" 
dataset <- tourism 
```

```{r, eval=TRUE}
sel_Data <- selectedData(dataset = dataset, state = state, region = region)

plot1 <- plotTs(filtered_data = sel_Data)
sR <- summarizeRegion(dataset = dataset, state = state)

head(sel_Data) # filtered data
```

## Testing 1) - answer - South Australia as state and Adelaide as region

```{r, eval=TRUE}
plot1 # time series plot
```

## Testing 1) - answer - South Australia as state and Adelaide as region

```{r, eval=TRUE}
sR # number of trips by region (dependent on the filtered state)
```

## Testing 2) - answer - South Australia as state and Adelaide as region and Holiday as purpose


```{r, eval=TRUE}
purpose <- "Holiday"
sel_Data_purpose <- selectedData(dataset = dataset, state = state, 
                                 region = region, purpose = purpose)

plot_purpose <- plotTs(filtered_data = sel_Data_purpose)
sR2 <- summarizeRegion(dataset = dataset, state = state)

head(sel_Data_purpose)
```


## Testing 2) - answer - South Australia as state and Adelaide as region and Holiday as purpose

```{r, eval=TRUE}
plot_purpose # time series plot
```

## Testing 2) - answer - South Australia as state and Adelaide as region and Holiday as purpose

```{r, eval=TRUE}
# no difference expected since the number of trips has been summarized by region 
# (is independent of purpose):
unique(sR == sR2) 
```

## Task - Maintaining

### Your turn

You have shown your program to a colleague. She/he recommends to compute also the percentage of trips of a selected Australian state. Create a function named `summarizePercTrips()`.


## Maintaining - answer

```{r}
summarizePercTrips <- function(dataset, state){
   sum_T <- dataset %>% 
   filter(State == state) %>%
    summarize(Trips = round(sum(Trips))) # sum of trips of the chosen state
   
   sum_T_total <- dataset %>% # total sum of trips (of all states)
     summarize(Trips = sum(Trips))
   
   rel_num_Trips <- sum_T / sum_T_total # sum of trips of chosen state / total sum
    return(rel_num_Trips)
}
```


## What can be improved?

So far, we have implemented a program with some functions in order to extract information we are interested in from the Australian tourism data. To make our program more user-friendly, it is desired

- to have a user interface, where we can select a state, region and purpose

    - to show all possible regions of a selected state
    
    - to show all purposes after selecting a state and region

- that the time series plot will automatically be updated when the input changes

- ...

$\rightarrow$ All in all, we want to examine the tourism data set interactively.

\begin{textblock}{11}(1,7.5)
\begin{alertblock}{}
A smart solution that allows these features is a \textbf{Shiny Web App} (Chang et al. (2021)) as we will see in a future exercise!
\end{alertblock}
\end{textblock}

## For home

Document the function `summarizePercTrips()`.

## References
