---
title: "Programmierung R - Exercise"
author: "Vectorization"
date: "May 10, SS 2022 || Hannah Behrens"

toc: true
tocoverview: false
tocheader: Inhaltsübersicht
titlefontsize: 22pt

bibliography: C:/Users/Mitarbeiterin/sciebo/nongit/SoSe_22/cswr_R_Programmierung_2022/Uebungen_Slides/referencescswr.bib
biblio-style: authoryear-comp

output: fhswf::presentation
knit: rmarkdown::render

nocite: |
    @cswr
---

```{r setup, include=FALSE}
library(knitr)
options(digits = 4)
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, cache=TRUE, 
                      dev.args=list(pointsize=11), size="scriptsize", fig.height = 4, attr.source='.numberLines')

# Packages
library(car)
library(fGarch)
library(gtools)
library(HistogramTools)
library(ineq)
library(kableExtra)
library(texreg)
library(tidyverse)
library(vcd)
library(latex2exp)
library(knitcitations)
library(gridExtra)
library(plotly)
library(rgl)
library(mgcv)
library(parallel)
```

## Task 1

### Your turn

1.1 Load the data set `Airquality3.csv` we have created at our second appointment from April 12, 2022 (Exercise *Data manipulation*).

1.2  Compute the mean of each numeric variable (use different `apply`- functions (`apply()`, `lapply()`, `sapply()` and `vapply()`)). How do the outputs differ?

1.3 Compute the median of `Temperature` dependent on the month.


## Tasks 1.1 and 1.2 - answers

1.1 
```{r, echo=FALSE, eval=TRUE}
# read csv-file named "Airquality3.csv" (exercise from the 12th of April 2022)
gwd <- getwd()
```

```{r}
# read csv-file named "Airquality3.csv" (exercise from April 12, 2022)
airqu3 <- read.table("Airquality3.csv", sep = ",")
```

1.2 

```{r}
# the first six variables are numeric:
str(airqu3)
```

## Task 1.2 - `apply()` - answer


```{r}
# apply the mean function to the first six columns of airqu3
airqu3_apply <- apply(X = airqu3[,c(1:6)], MARGIN = 2, FUN = mean) 
airqu3_apply
typeof(airqu3_apply) # output: a vector
```

\begin{textblock}{4}(6,1)
\begin{alertblock}{}
function \texttt{apply()}
\end{alertblock}
\end{textblock}

## Task 1.2 - `lapply()` - answer

```{r}
airqu3_lapply <- lapply(X = airqu3[,c(1:6)], FUN = mean)
airqu3_lapply
typeof(airqu3_lapply) # output: a list
```

\begin{textblock}{4}(6,4)
\begin{alertblock}{}
function \texttt{lapply()},\newline
return value: a list
\end{alertblock}
\end{textblock}


## Task 1.2 - answer

```{r}
air_list <- list(Ozone = airqu3$Ozone, Solar.R = airqu3$Solar.R,
                 Wind = airqu3$Wind, Temperature = airqu3$Temperature,
                 Month = airqu3$Month, Day = airqu3$Day)
air_list
```


\begin{textblock}{6}(5,2)
\begin{alertblock}{}
Saving \texttt{airqu3} as a list named \texttt{air\_list}
\end{alertblock}
\end{textblock}

## Task 1.2 - `lapply()` - answer

```{r}
airqu3_lapply2 <- lapply(X = air_list, FUN = mean)
airqu3_lapply2
typeof(airqu3_lapply2) # output: a list
```

\begin{textblock}{7}(4,3)
\begin{alertblock}{}
applying \texttt{lapply()} to a list $\rightarrow$ a list is returned
\end{alertblock}
\end{textblock}

## Task 1.2 - `sapply()` - answer


```{r}
airqu3_sapply <- sapply(X = airqu3[,c(1:6)], FUN = mean)
airqu3_sapply
typeof(airqu3_sapply) # output: a vector
```

\begin{textblock}{4}(6,1)
\begin{alertblock}{}
function \texttt{sapply()}
\end{alertblock}
\end{textblock}

## Task 1.2 - `sapply()` - answer

```{r}
airqu3_sapply2 <- sapply(air_list, FUN = mean)
airqu3_sapply2
typeof(airqu3_sapply2) # output: a vector
```

\begin{textblock}{11}(1,1)
\begin{alertblock}{}
applying \texttt{sapply()} to a list $\rightarrow$ a vector is returned when \texttt{simplify = TRUE}
\end{alertblock}
\end{textblock}

\begin{textblock}{7}(2,7)
\begin{alertblock}{}
What happens when \texttt{simplify} is set to \texttt{FALSE}?
\end{alertblock}
\end{textblock}

## Task 1.2 - `sapply()` - answer

```{r}
airqu3_sapply3 <- sapply(air_list, FUN = mean, simplify = FALSE)
airqu3_sapply3
typeof(airqu3_sapply3) # output: a vector
```

\begin{textblock}{7}(4,2)
\begin{alertblock}{}
Setting \texttt{simplify = FALSE}: a list is returned
\end{alertblock}
\end{textblock}

## Task 1.2 - `vapply()` - answer

```{r}
airqu3_vapply <- vapply(X = airqu3[,c(1:6)], FUN = mean, FUN.VALUE = numeric(1))
airqu3_vapply
typeof(airqu3_vapply) # output: a vector
```

\begin{textblock}{11}(0.7,1)
\begin{alertblock}{}
function \texttt{vapply()}: specification of the return value (\texttt{FUN.VALUE}) is necessary
\end{alertblock}
\end{textblock}


## Task 1.3 - answer

1.3 Based on `Airquality3.csv`, compute the median of `Temperature` dependent on the month.

```{r}
tapply(X = airqu3$Temperature, INDEX = airqu3$Month, FUN = median)
# or
tapply(X = airqu3$Temperature, INDEX = airqu3$Month_name, FUN = median)

# --> is a big advantage compared to calculating the median of the temperature 
# for each month piecewise like
median(airqu3$Temperature[which(airqu3$Month == 5)]) 
# ... and so on
```


## Task 2

Remember the data set `datasets::airquality` [@datasets]. It consists of the following variables

```{r}
colnames(airquality)
```

which are numeric.

We are interested in all possible and different linear models when one of these six variables is considered as the dependent variable $y$. How can we achieve this?

\begin{textblock}{10}(0.7, 8)
\begin{alertblock}{}
How many combinations of different linear models exist for each $y$?
\end{alertblock}
\end{textblock}

## Task 2 - number of combinations - answer

For each $y$: 

+ $N = 5$ remaining variables as predictors to form linear models, e.g. $y$ = `Ozone`:

```{r, eval=FALSE}
lm(Ozone ~ 1 + Solar.R + Wind + Temp + Month + Day)
# but also
lm(Ozone ~ 1 + Solar.R + Wind + Temp + Month)
# as well as
lm(Ozone ~ 1 + Solar.R + Wind + Temp + Day)
# and so on...
```


+ As we have seen, a linear model for `Ozone` can be created based on 5, 4, 3, 2 and 1 variable(s) as predictor(s). 

+ Each predictor appears only **once** in each linear model and the ordering **does not** play any role: So, the number of combinations
 when picking up 5 of 5, 4 of 5 predictors and so on is calculated based on the binomial coefficient $\binom{N}{n} = \frac{N!}{n! \cdot (N-n)!}$. [@Fahrmeir2016, p. 187f.]

## Task 2 - number of combinations - answer

From that it follows:

+ When forming a model with 5 predictors, only 1 combination is possible: $\binom{5}{5} = \frac{5!}{5! \cdot (5-5)!} = 1$

+ When forming a model with 4 predictors, 5 combinations are possible: $\binom{5}{4} = \frac{5!}{4! \cdot (5-4)!} = 5$

$\rightarrow$ So, for each $y$ 

```{r}
combi_per_y <- choose(5,5) + choose(5,4) + choose(5,3) + choose(5,2) + choose(5,1)
combi_per_y
```

different linear models can be created and examined.


## Task 2 - implementation - step 1

### Your turn 

Based on the presented theory, the first step is to *combine the predictors* to create `r combi_per_y` linear models in the next step. These *combinations* should be saved in a list.\newline
How can we do this?


## Task 2 - implementation - step 1 - answer


```{r, eval=TRUE}
## main program
data <- airquality
coln_names <- colnames(data) # get column names

y <- coln_names[1] # e.g. Ozone (column 1) as the dependent variable (y)
vars_grid <- coln_names[-1] # the remaining variables are predictors
variables_grid <- expand.grid(replicate(length(vars_grid), vars_grid,
                     simplify = FALSE)) # make a grid with all possible combinations
    # since each predictor can only be listed once in the linear model, remove replicates
variables_grid_unique <- apply(variables_grid, MARGIN = 1, FUN = unique)
# sort entries of the sublists since the order of the predictors does not play any role
variables_grid_sort <- lapply(variables_grid_unique, sort)
    # then remove replicates again
variables_grid_sort_unique <- unique(variables_grid_sort)
   # add y to the end of all sublists:
variables_grid_sort_unique_y <- lapply(variables_grid_sort_unique, 
                                FUN = function(x){c(x, x[length(x)+1] <- y)})
air_grid_lm <- variables_grid_sort_unique_y
    # ... something is missing here
```

\begin{textblock}{10}(0.7, 8)
\begin{alertblock}{}
What to do next? Any ideas? Suggestions?
\end{alertblock}
\end{textblock}

\begin{textblock}{8}(4, 0.8)
\begin{alertblock}{}
Combining the predictors to create `r combi_per_y` linear models
\end{alertblock}
\end{textblock}

## Task 2 - step 2 

### Your turn

Based on our list `air_grid_lm`, we have to create the formulas of each linear model with $y$ (here: `Ozone`) as dependent variable. How can you implement it? Do it!

## Task 2 - step 2 - implementation - answer



```{r, echo=FALSE, eval=TRUE}
gwd <- getwd()
source("determiningLinMod.R")
```


```{r, eval=FALSE}
# x: the first element(s) of each sublist (= predictors)
# y: the last element of each sublist
determiningLinMod <- function(x, y = x[length(x)], dataset = airquality){
  # predictors x1 ,..., x_{length(x)-1}
  predictors <- paste(x[-length(x)], collapse = " + ") # concatenate the predictors 
  # by a + -sign
  make_formula <- paste(y, " ~ 1 + ", predictors)
  formula_lm <- as.formula(make_formula) # convert the string to a formula
  lin_model <- lm(formula_lm, data = dataset) # call lm()
  # return the different formulas as well as the model
  return(list(formula = formula_lm, model = lin_model))
}
```

```{r, eval=TRUE}
# apply the function determiningLinMod() to each combination listed in air_grid_lm:
lapply_lm_air <- lapply(X = air_grid_lm, FUN = determiningLinMod)
```

## Task 2 - step 2 - implementation - answer

```{r, eval=FALSE}
## main program
data <- airquality
coln_names <- colnames(data) # get column names

y <- coln_names[1] # e.g. Ozone (column 1) as the dependent variable (y)
vars_grid <- coln_names[-1] # the remaining variables are predictors
variables_grid <- expand.grid(replicate(length(vars_grid), vars_grid,
                     simplify = FALSE)) # make a grid with all possible combinations
    # since each predictor can only be listed once in the linear model, remove replicates
variables_grid_unique <- apply(variables_grid, MARGIN = 1, FUN = unique)
# sort entries of the sublists since the order of the predictors does not play any role
variables_grid_sort <- lapply(variables_grid_unique, sort)
    # then remove replicates again
variables_grid_sort_unique <- unique(variables_grid_sort)
   # add y to the end of all sublists:
variables_grid_sort_unique_y <- lapply(variables_grid_sort_unique, 
                                FUN = function(x){c(x, x[length(x)+1] <- y)})
air_grid_lm <- variables_grid_sort_unique_y
# insert:
lapply_lm_air <- lapply(air_grid_lm, FUN = determiningLinMod)
```

## Task 2 - step 2 - implementation - results

```{r}
length(lapply_lm_air) # 31 combinations expected (see some slides before)


lapply_lm_air
```


## Task 2 - step 3 - parallelization

Very nice! We get all possible linear models for the variable `Ozone` of the `airquality` data set [@datasets].

But now, we want to go a step further by parallelizing the determination of the `r combi_per_y`  linear models for `Ozone`.

\begin{textblock}{4}(4, 6.5)
\begin{alertblock}{}
How can we achieve this?
\end{alertblock}
\end{textblock}

## Task 2 - step 3 - parallelization

The function `parallel::mclapply()` [@parallel] is available which works as follows (The example is taken from @Peng2020):


```{r}
 r <- mclapply(1:10, function(i) {
         Sys.sleep(10)  # Do nothing for 10 seconds
}, mc.cores = 1)
r
```


## Task 2 - step 3 - parallelization

\begin{textblock}{6}(6, 2)
\begin{alertblock}{}
Instead of using \texttt{lapply()} \texttt{mclapply} (R Core Team 2021b) is used.\newline Parallelization based on \texttt{mclapply} is not possible on Windows.
\end{alertblock}
\end{textblock}

```{r}
parallelized_lm_air <- parallel::mclapply(X = air_grid_lm, 
                                          FUN = determiningLinMod, mc.cores = 1)
parallelized_lm_air
```


## Task 2 - step 3 - parallelization

```{r}
length(parallelized_lm_air)
```


\begin{textblock}{6}(4, 6.5)
\begin{alertblock}{}
The expectations have been fulfilled!
\end{alertblock}
\end{textblock}

## Task 3


The following list is given:

```{r}
L <- list(money = c(250, 124, 360, 720, 340, 340),
          hours = c(19, 12, 30, 48, 26, 25),
          idx = c(1:6),
          name = c("Paul", "Emma", "Mia", "John", "Kim", "Maxi"))
```

### Your turn
Make use of the `*apply()`-functions and

a) compute the total sum of `money`, `hours` and `idx` and return a list.

b) compute the mean of `money` and `hours` and return a vector.

c) multiply each element of `money` and `hours` by 1.5 and return a matrix.

## Tasks 3a) and 3b) - answers

```{r}
# 3.a)
lapply(X = L[1:3], FUN = sum)

# 3.b) 
sapply(X = L[1:2], FUN = mean)
# vapply(L[1:2], mean, FUN.VALUE=numeric(1)) # alternative
# --> the value specified by FUN.VALUE (here: numeric(1)) and the actual value returned
# (here: the mean of money and hours) have to match
```


## Task 3.c) - answer

```{r}
sapply(X = L[1:2], FUN = function(x){x * 1.5})
```

## For home

Document the code (main program (slide 23) and function \texttt{determiningLinMod()}) as you have learned in the last lecture and exercise about *Software Development*.\newline
You can also think of wrapping the code from slide 23 in a function.

## References